---
title: "Career Readiness: What Students See vs. What Supervisors See"
subtitle: "Debrief: Career Readiness Survey Analysis"
author: "Antoine Ward, MPH"
date: today
format:
  revealjs:
    slide-number: true
    footer: "Career Readiness Analytics: 22 NACE-aligned competencies, 7 dimensions"
    transition: fade
    incremental: false
  pptx: default            # `quarto render debrief.qmd --to pptx` for PowerPoint
engine: knitr
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(readr); library(dplyr); library(tidyr); library(ggplot2)

dim_summary <- read_csv("data/exports/gold_dimension_summary.csv",   show_col_types = FALSE)
gap         <- read_csv("data/exports/gold_perception_gap.csv",      show_col_types = FALSE)
comp        <- read_csv("data/exports/gold_competency_summary.csv",  show_col_types = FALSE)

C_STUDENT <- "#4C78A8"; C_OBSERVER <- "#F58518"
C_UNDER   <- "#54A24B"; C_OVER     <- "#E45756"
```

## The question we asked {.smaller}

**Do students and their workplace supervisors see student career readiness the same way?**

- ~**840 students** rated themselves on 22 competencies
- ~**1,000 supervisors** rated students on the *identical* 22
- Competencies roll up to **7 NACE career readiness dimensions**
- Rating scale: **1–4**

::: {.fragment}
Same questions, two vantage points, so every disagreement is a *perception gap* we can measure.
:::

## The headline finding

::: {.r-fit-text}
Students under-rate themselves

on **all seven** dimensions.
:::

::: {.fragment}
Not overconfidence: a **confidence gap**. The opposite of what the conventional wisdom predicts.
:::

## Self vs. supervisor, by dimension

```{r dumbbell, fig.width=9, fig.height=5.2}
wide <- dim_summary %>%
  select(dimension, respondent_type, mean_score) %>%
  pivot_wider(names_from = respondent_type, values_from = mean_score) %>%
  mutate(dimension = reorder(dimension, (student + observer) / 2))

ggplot(wide) +
  geom_segment(aes(x = observer, xend = student, y = dimension, yend = dimension),
               color = "grey70", linewidth = 1.2) +
  geom_point(aes(x = student,  y = dimension, color = "Student (self)"), size = 5) +
  geom_point(aes(x = observer, y = dimension, color = "Supervisor"),     size = 5) +
  scale_color_manual(values = c("Student (self)" = C_STUDENT, "Supervisor" = C_OBSERVER)) +
  scale_x_continuous(limits = c(2.5, 4), breaks = seq(2.5, 4, 0.5)) +
  labs(x = "Mean score (1\u20134 scale)", y = NULL, color = NULL,
       title = "Supervisors rate students higher on every dimension") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "top")
```

## The full ranking {.smaller}

Strongest to weakest, side by side

```{r ranktable}
rank_tbl <- dim_summary %>%
  group_by(respondent_type) %>%
  mutate(rank = min_rank(desc(mean_score))) %>%
  ungroup() %>%
  select(dimension, respondent_type, mean_score, rank) %>%
  pivot_wider(names_from = respondent_type, values_from = c(mean_score, rank)) %>%
  arrange(rank_student) %>%
  select(Dimension = dimension,
         `Student mean` = mean_score_student, `Student rank` = rank_student,
         `Supervisor mean` = mean_score_observer, `Supervisor rank` = rank_observer)
knitr::kable(rank_tbl, align = "lcccc")
```

::: {.fragment}
The *order* barely changes between raters. Professionalism and Teamwork sit at the top for
both; Leadership and Career & Self-Development sit at the bottom.
:::

## Where the gap is largest

```{r gapchart, fig.width=9, fig.height=5.2}
gap %>%
  mutate(dimension = reorder(dimension, gap)) %>%
  ggplot(aes(x = gap, y = dimension)) +
  geom_col(fill = C_UNDER) +
  geom_vline(xintercept = 0) +
  geom_text(aes(label = sprintf("%+.2f", gap)), hjust = 1.15, size = 4.5, color = "white") +
  labs(x = "Gap  (student mean \u2212 supervisor mean)", y = NULL,
       title = "Every bar points the same way: students under-rate",
       subtitle = "Largest under-estimates: Career & Self-Development, Communication") +
  theme_minimal(base_size = 15)
```

## The numbers behind the gaps {.smaller}

```{r gaptable}
gap %>%
  arrange(gap) %>%
  transmute(Dimension   = dimension,
            Student     = round(student_mean, 2),
            Supervisor  = round(observer_mean, 2),
            Gap         = round(gap, 2),
            `Cohen's d` = round(cohens_d, 2)) %>%
  knitr::kable(align = "lcccc")
```

::: {.fragment}
Every row reads *student under-rates*: the pattern is perfectly consistent, even
where the gap itself is small.
:::

## Two different "weakest" dimensions {.smaller}

| Vantage point | Weakest dimension | Score |
|---|---|---|
| Students (self) | **Career & Self-Development** | 2.92 / 4 |
| Supervisors | **Leadership** | 3.14 / 4 |

::: {.fragment}
That divergence *is* the insight:

- **Leadership** is weakest by *external* assessment: the clear **skill-development** target
- **Career & Self-Development** combines students' lowest self-rating *and* their largest under-estimate (−0.24): a **confidence & awareness** target, not just a skill one
:::

## Inside the dimensions: all 22 competencies {.smaller}

```{r heatmap, fig.width=8.5, fig.height=6.2}
comp %>%
  mutate(rater = ifelse(respondent_type == "student", "Student", "Supervisor"),
         competency = reorder(competency, mean_score)) %>%
  ggplot(aes(x = rater, y = competency, fill = mean_score)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", mean_score)), size = 2.8) +
  scale_fill_gradient2(low = C_OVER, mid = "#F7E9A0", high = C_UNDER,
                       midpoint = mean(comp$mean_score), name = "Mean") +
  facet_grid(dimension ~ ., scales = "free_y", space = "free_y",
             labeller = label_wrap_gen(width = 14)) +
  labs(x = NULL, y = NULL,
       title = "Competency-level detail behind the dimension averages") +
  theme_minimal(base_size = 11) +
  theme(strip.text.y = element_text(angle = 0, size = 7, hjust = 0),
        legend.position = "right",
        axis.text.y = element_text(size = 7.5))
```

## What the heatmap adds {.smaller}

Dimension averages can hide uneven competencies. Three details worth naming:

- **Act with Integrity** and **Demonstrate Dependability** are the highest-rated
  competencies overall: the engine behind Professionalism's top ranking.
- Within Leadership, the weakness is not uniform: *Inspire, Persuade & Motivate*
  sits lower than *Engage Resources & Seek Feedback*, so leadership programming
  should target influence skills specifically, not leadership generically.
- **Networking** is among students' lowest self-ratings, consistent with the
  Career & Self-Development confidence story.

## Individuals vary more than groups differ

```{r spread, fig.width=9, fig.height=5.2}
dim_summary %>%
  mutate(rater = ifelse(respondent_type == "student", "Student (self)", "Supervisor")) %>%
  ggplot(aes(x = mean_score, y = reorder(dimension, mean_score), color = rater)) +
  geom_pointrange(aes(xmin = mean_score - sd_score, xmax = mean_score + sd_score),
                  position = position_dodge(width = 0.55), linewidth = 0.9, size = 0.6) +
  scale_color_manual(values = c("Student (self)" = C_STUDENT, "Supervisor" = C_OBSERVER)) +
  scale_x_continuous(limits = c(1, 4), breaks = 1:4) +
  labs(x = "Mean score \u00b1 1 standard deviation (1\u20134 scale)", y = NULL, color = NULL,
       title = "The spread within each group dwarfs the gap between groups",
       subtitle = "Each bar spans \u00b11 SD: the range where most individual ratings fall") +
  theme_minimal(base_size = 15) +
  theme(legend.position = "top")
```

## How big are these gaps, honestly? {.smaller}

**Cohen's d** answers one question: *"Is this gap large enough to matter, or
just noise?"* It rescales the raw gap by how spread out people's scores
naturally are: the spread you just saw. A 0.24-point gap on a tight,
consistent dimension counts for more than the same 0.24 points where scores
bounce around.

::: {.fragment}
**Rule of thumb:**

| \|d\| | Interpretation |
|---|---|
| ~0.2 | small (barely noticeable) |
| ~0.5 | medium (visible in practice) |
| ~0.8+ | large (hard to miss) |
:::

::: {.fragment}
Our largest gap (Career & Self-Development) is **d = −0.29** (small).
**No dimension reaches 0.5.**

**Translation:** students and supervisors are *broadly calibrated*. These gaps
are **monitor items, not crises**, but their perfect directional consistency
(7 of 7 negative) is a real pattern worth acting on regardless of size.
:::

## One agreement worth noting

```{python professionalism}
# Python via reticulate: competency detail behind the shared top strength
import pandas as pd
dim = pd.read_csv("data/exports/gold_dimension_summary.csv")
prof = (dim[dim["dimension"] == "Professionalism"]
        [["respondent_type", "mean_score", "n_respondents"]]
        .rename(columns={"respondent_type": "Rated by",
                         "mean_score": "Mean (1-4)",
                         "n_respondents": "N"}))
prof["Rated by"] = prof["Rated by"].str.title()
print(prof.to_string(index=False))
```

Both groups rank **Professionalism** as the top strength (3.41 self, 3.52 supervisor). Agreement at the top says the assessment is coherent, which strengthens confidence in the gaps.

## Recommendations {.smaller}

::: {.incremental}
1. **Build skills where supervisors see the need: Leadership** (3.14/4).
   Prioritize facilitation, delegation, and peer-leadership practice inside
   experiential placements (e.g., rotate students through a project-lead
   role for at least one deliverable per placement).
2. **Build confidence where students underclaim: Career & Self-Development.**
   Students demonstrate more competence than they recognize; add structured
   self-assessment checkpoints and strengths-based reflection at the
   midpoint and end of each placement.
3. **Close the calibration loop: structured supervisor feedback.**
   The uniform under-rating suggests students rarely *hear* how supervisors
   rate them. Build a short mid-placement feedback conversation into every
   experiential learning opportunity; even 15 minutes moves the needle.
4. **Track this survey longitudinally.** Re-run it each term to see whether
   the gap narrows as feedback practices take hold, turning this one-time
   snapshot into a measure of program impact.
:::

## Methodology & caveats {.smaller}

- **Pipeline:** medallion lakehouse (Bronze → Silver → Gold) on Databricks; analysis
  reproducible in SQL, PySpark, and R
- **Unit of analysis:** dimension scores averaged *within respondent first*; effect
  sizes count people, not survey items
- **Unpaired comparison:** student and supervisor pools, not matched pairs; gaps
  describe *group-level* divergence, not individual miscalibration
- **Cross-sectional:** associations, not causal claims
- **Scale note:** Equity & Inclusion (8th NACE competency) was not surveyed

## Questions

**Code, data pipeline, and full write-up:**
github.com/antoinewrd1/career-readiness-analytics

*Analysis: Databricks (PySpark, Spark SQL, Delta Lake) · R (dplyr, ggplot2) ·
Python (pandas, matplotlib) · Power BI*
